### 2017-03-31 paper/oopsla17-submission/overview.tex
minor

\paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
We start by assuming that we will extract features from the program's
abstract syntax tree (AST), a natural choice since the type checker
operates on ASTs.
%
An obvious set of interesting features is then the syntactic class of
each expression, \ie the classifier should be able to distinguish
between the |[]| and |+| expressions in \autoref{fig:sumList} because
they are formed by different constructors in the AST.
%
A programming language has a fixed set of AST elements, which suggests a
natural way to obtain fixed-size feature vectors: rather than translating
a program in a single feature vector, we translate it into a \emph{set} of
feature vectors, one for each expression in the program.
%
Thus, the |[]| and |+| expressions will each be given their own feature
vector, and we will ask the classifier to predict for each
\emph{independently} whether it should be blamed.

@@

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix, but this approach does not
scale well to a dataset large enough to perform machine learning.
%
Furthermore, while the expert user has intimate knowledge of the type
system, she cannot know what the student \emph{intended} to do.

\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
%
Rather than relying on expert users, we allow our students to do the
work of assigning blame themselves.
%
Software development is an iterative process and our students will
eventually fix their own programs, perhaps after multiple ill-fated
attempts.
%
For each ill-typed program in our dataset, we find the first
\emph{subsequent} program that type checks and declare that to be the
fix.
%
From this pair of an ill-typed program and its fix, we can extract a
\emph{diff} of the abstract syntax trees.
%
For example, suppose our student fixed the |sumList| program in
\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
only the |[]| expression.
%
Thus we would determine that the |[]| expression (and \emph{not} the
|+|) is to blame for the error.

\paragraph{\textbf{Challenge (3): What should be presented to the user?}}
We now have the tools to train a classifier to predict, for each
expression in an ill-typed program, whether it should be changed.
%
But this is not yet particularly suitable as \emph{user feedback}.
%
A recent survey of developers by \citet{Kochhar2016-oc} found that
developers are unlikely to examine more than around five potentially
erroneous locations before falling back to manual debugging.
%
Thus, we should limit our predictions to a select few to be presented to
the user.

\paragraph{\textbf{Solution: Rank expressions by ``confidence''}}
Luckily, many machine learning classifiers produce not only a predicted
label, but also a metric that can be interpreted as the classifier's
\emph{confidence} in its prediction.
%
Thus, we \emph{rank} each expression by the classifier's confidence that
it should be changed, and present only the top-$k$ predictions to the
user (in practice $k=3$).

>>>>

\paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
\ES{name up for grabs...}
We start by assuming that we will extract features from the program's
abstract syntax tree (AST), a natural choice since the type checker
operates on ASTs.
%
Features would then be \emph{queries} over the AST, an obvious set of
interesting features to start with is the syntactic class of each
expression.
%
That is, the classifier should be able to distinguish between the |[]|
and |+| expressions in \autoref{fig:sumList} because they represent
different types of expressions.
%
A programming language has a fixed set of AST elements\footnote{We will
  discuss handling arbitrary, user-defined types in
  \autoref{sec:discussion}.}, which suggests a natural way to obtain
fixed-size feature vectors: rather than translating a program in a
single feature vector, we translate it into a \emph{set} of feature
vectors, one for each expression in the program.
%
Thus, the |[]| and |+| expressions will each be given their own feature
vector, and we will ask the classifier to predict for each
\emph{independently} whether it should be blamed.

@@

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has often enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix, but this approach does not
scale well to a dataset large enough to perform machine learning.
%
Furthermore, while the expert user has intimate knowledge of the type
system, she cannot in general know what the student \emph{intended} to do.

\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
%
Rather than relying on expert users, we allow our students to do the
work of assigning blame themselves.
%
Software development is an iterative process and our students will
eventually fix their own programs, perhaps after multiple ill-fated
attempts.
%
For each ill-typed program in our dataset, we find the first
\emph{subsequent} program that type checks and declare that to be the
fix.
%
From this pair of an ill-typed program and its fix, we can extract a
\emph{diff} of the abstract syntax trees.
%
For example, suppose our student fixed the |sumList| program in
\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
only the |[]| expression.
%
Thus we would determine that the |[]| expression (and \emph{not} the
|+| or the recursive call |sumList t|) is to blame for the error.

\paragraph{\textbf{Challenge (3): What should be presented to the user?}}
We now have the features and labels needed to train a classifier to
predict, for each expression in an ill-typed program, whether it should
be blamed.
%
But this is not yet particularly suitable as \emph{user feedback}.
%
A recent survey of developers by \citet{Kochhar2016-oc} found that
developers are unlikely to examine more than around five potentially
erroneous locations before falling back to manual debugging.
%
Thus, we should limit our predictions to a select few to be presented to
the user.

\paragraph{\textbf{Solution: Rank blame locations by ``confidence''}}
Luckily, many machine learning classifiers produce not only a predicted
label, but also a metric that can be interpreted as the classifier's
\emph{confidence} in its prediction.
%
Thus, we \emph{rank} each expression by the classifier's confidence that
it should be blamed, and present only the top-$k$ predictions to the
user (in practice $k=3$).
